---
title: "PHTS_mt_CN_analysis"
author: "R.W."
output: html_document
---

# knit setting
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library
```{r}
library(ggplot2)
library(ggsignif)
library(ggbeeswarm)
library(ggpubr)
library(beeswarm)
library(vipor)
library(ggfortify)
library(VennDiagram)
library(RColorBrewer)
library(corrplot)
library(factoextra)
library(scales)
library(gridExtra)
```

# read in all polished data
```{r}
all = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_all.txt",sep="\t")
short_phenotype = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_short_phenotypes_wCN.csv",sep="\t")
short_all = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_short_all.csv",sep="\t")
```

# phenotype annotation 
```{r}
onlyASD = short_all[which(short_all$phenotype=="ASD"),]
onlyASD$phenotype = "onlyASD"
onlyCancer = short_all[which(short_all$phenotype=="Cancer"),]
onlyCancer$phenotype = "onlyCancer"
ASDCancer = short_all[which(short_all$phenotype=="ASDCancer"),]
neither = short_all[which(short_all$phenotype=="neither"),]
ASD = rbind.data.frame(onlyASD,ASDCancer)
Cancer = rbind.data.frame(onlyCancer,ASDCancer)
noASD = rbind.data.frame(onlyCancer,neither)
noCancer = rbind.data.frame(onlyASD,neither)

for_CN=rbind.data.frame(onlyASD,onlyCancer,ASDCancer,neither)
for_CN$AGCNST = as.numeric(for_CN$AGCNST)
```

# beeswarm plot of CN distribution of within every haplogroup
```{r}
ggplot(data=short_phenotype,aes(y=CN,x=haplogroup)) +
  #geom_histogram(alpha=0.6) +
  #geom_density(fill="lightblue") +
  geom_boxplot() + 
  geom_point(alpha=0.6,aes(col=haplogroup),position=position_jitter(w=0.4)) +
  #ggtitle("") +
  theme_bw() +
  ylab("mtDNA CN") +
  #scale_color_brewer(palette="Set3") +
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(size=15),
        legend.position = "none",
        title = element_text(size=20))
```

# beeswarm plot of CN distribution between H and non-H haplogroups
```{r}
short_phenotype$H = "non-H"
short_phenotype[which(short_phenotype$haplogroup=="H"),"H"] = "H"

ggplot(data=short_phenotype,aes(y=CN,x=H)) +
  #geom_histogram(alpha=0.6) +
  #geom_density(fill="lightblue") +
  geom_boxplot() + 
  geom_point(alpha=0.6,aes(col=H),position=position_jitter(w=0.4)) +
  #ggtitle("Copy Number Density Plot") +
  ylab("mtDNA CN") +
  xlab("haplogroup") +
  theme_bw() +
  scale_color_brewer(palette="Accent") +
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(size=15),
        legend.position = "none",
        title = element_text(size=20))
```
## MW test
```{r}
wilcox.test(short_phenotype[which(short_phenotype$H=="H"),"CN"],short_phenotype[which(short_phenotype$H=="non-H"),"CN"])
```

# CN MW test for all sample
## onlyASD, onlyCancer, ASDCancer, neither
```{r}
t1 = wilcox.test(onlyASD$CN,onlyCancer$CN)[3]
t2 = wilcox.test(onlyASD$CN,ASDCancer$CN)[3]
t3 = wilcox.test(ASDCancer$CN,onlyCancer$CN)[3]
t4 = wilcox.test(onlyASD$CN,neither$CN)[3]
t5 = wilcox.test(onlyCancer$CN,neither$CN)[3]
t6 = wilcox.test(ASDCancer$CN,neither$CN)[3]

CN_heat = rbind.data.frame(
c(t1,"onlyASD","onlyCancer"),
#c(t2,"onlyASD","ASDCancer"),
c(t3,"ASDCancer","onlyCancer"),
#c(t4,"onlyASD","neither"),
#c(t5,"onlyCancer","neither"),
c(t6,"ASDCancer","neither"),
#c(t1,"onlyCancer","onlyASD"),
c(t2,"ASDCancer","onlyASD"),
#c(t3,"onlyCancer","ASDCancer"),
c(t4,"neither","onlyASD"),
c(t5,"neither","onlyCancer"),
#c(t6,"neither","ASDCancer"),
c(1,"ASDCancer","ASDCancer"),
c(1,"onlyASD","onlyASD"),
c(1,"onlyCancer","onlyCancer"),
c(1,"neither","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)
```

## ASD, no ASD, neither
```{r}
ASD$asd_phenotype = "ASD"
noASD$asd_phenotype = "noASD"
neither$asd_phenotype = "neither"
for_CN_ASD=rbind.data.frame(ASD,noASD)
for_CN_ASD$AGCNST = as.numeric(for_CN_ASD$AGCNST)
for_CN_ASD$asd_phenotype = factor(for_CN_ASD$asd_phenotype,levels = c("ASD","noASD","neither"))
```

```{r}
t1 = wilcox.test(ASD$CN,noASD$CN)[3]
t2 = wilcox.test(ASD$CN,neither$CN)[3]
t3 = wilcox.test(neither$CN,noASD$CN)[3]


CN_heat_ASD = rbind.data.frame(
c(t1,"ASD","noASD"),
c(t2,"ASD","neither"),
c(t3,"neither","noASD"),
c(1,"ASD","ASD"),
c(1,"noASD","noASD"),
c(1,"neither","neither"))

colnames(CN_heat_ASD) = c("p value","phenotype1","phenotype2")
CN_heat_ASD$`p value` = as.numeric(CN_heat_ASD$`p value`)
CN_heat_ASD$groups <- cut(CN_heat_ASD$`p value`,              
                       breaks = c(1,0.1,0.05,0))
```

```{r}
t1 = wilcox.test(ASD$CN,Cancer$CN)[3]
t2 = wilcox.test(ASD$CN,noASD$CN)[3]
t3 = wilcox.test(Cancer$CN,noCancer$CN)[3]
t4 = wilcox.test(Cancer$CN,neither$CN)[3]
t5 = wilcox.test(ASD$CN,neither$CN)[3]
t6 = wilcox.test(noASD$CN,neither$CN)[3]
t7 = wilcox.test(noCancer$CN,neither$CN)[3]

CN_heat = rbind.data.frame(
c(t1,"ASD","Cancer"),
c(t2,"ASD","noASD"),
c(t3,"Cancer","noCancer"),
c(t4,"Cancer","neither"),
c(t5,"ASD","neither"),
c(t6,"noASD","neither"),
c(t7,"noCancer","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)
```

## cancer, no cancer, neither
```{r}
Cancer$asd_phenotype = "NA"
noCancer$asd_phenotype = "NA"
Cancer$cancer_phenotype = "Cancer"
noCancer$cancer_phenotype = "noCancer"
neither$cancer_phenotype = "neither"
for_CN_cancer=rbind.data.frame(Cancer,noCancer)
for_CN_cancer$AGCNST = as.numeric(for_CN_cancer$AGCNST)
for_CN_cancer$cancer_phenotype = factor(for_CN_cancer$cancer_phenotype,levels = c("Cancer","noCancer","neither"))
```

```{r}
t1 = wilcox.test(Cancer$CN,noCancer$CN)[3]
t2 = wilcox.test(Cancer$CN,neither$CN)[3]
t3 = wilcox.test(neither$CN,noCancer$CN)[3]


CN_heat_cancer = rbind.data.frame(
c(t1,"Cancer","noCancer"),
c(t2,"Cancer","neither"),
c(t3,"neither","noCancer"),
c(1,"Cancer","Cancer"),
c(1,"noCancer","noCancer"),
c(1,"neither","neither"))

colnames(CN_heat_cancer) = c("p value","phenotype1","phenotype2")
CN_heat_cancer$`p value` = as.numeric(CN_heat_cancer$`p value`)

```
## ASD vs. Cancer
```{r}
ASD$cancer_phenotype = "NA"
noASD$cancer_phenotype = "NA"
onlyCancer$cancerasd_phenotype = "onlyCancer"
onlyCancer$asd_phenotype = "NA"
onlyCancer$cancer_phenotype = "NA"
onlyASD$asd_phenotype = "NA"
onlyASD$cancer_phenotype = "NA"
onlyASD$cancerasd_phenotype = "onlyASD"
ASD$cancerasd_phenotype = "ASD"
noASD$cancerasd_phenotype = "noASD"
Cancer$cancerasd_phenotype = "Cancer"
noCancer$cancerasd_phenotype = "noCancer"
neither$cancerasd_phenotype = "neither"
ASDCancer$asd_phenotype = "NA"
ASDCancer$cancer_phenotype = "NA"
ASDCancer$cancerasd_phenotype = "ASDCancer"
for_CN_cancerasd=rbind.data.frame(Cancer,noCancer,ASD,noASD,onlyASD,onlyCancer,neither,ASDCancer)
for_CN_cancerasd$AGCNST = as.numeric(for_CN_cancerasd$AGCNST)
for_CN_cancerasd$cancerasd_phenotype = factor(for_CN_cancerasd$cancerasd_phenotype,levels = c("onlyASD","ASD","noASD","onlyCancer","Cancer","noCancer","ASDCancer","neither"))
```

```{r}
t1 = wilcox.test(ASD$CN,noASD$CN)[3]
t2 = wilcox.test(ASD$CN,neither$CN)[3]
t3 = wilcox.test(ASD$CN,Cancer$CN)[3]
t4 = wilcox.test(onlyASD$CN,noASD$CN)[3]
t5 = wilcox.test(onlyASD$CN,neither$CN)[3]
t6 = wilcox.test(onlyASD$CN,onlyCancer$CN)[3]
t7 = wilcox.test(onlyCancer$CN,noCancer$CN)[3]
t8 = wilcox.test(onlyCancer$CN,neither$CN)[3]
t9 = wilcox.test(Cancer$CN,noCancer$CN)[3]
t10 = wilcox.test(Cancer$CN,neither$CN)[3]


CN_heat_cancerasd = rbind.data.frame(
c(t1,"ASD","noASD"),
c(t2,"ASD","neither"),
c(t3,"ASD","Cancer"),
c(t4,"onlyASD","noASD"),
c(t5,"onlyASD","neither"),
c(t6,"onlyASD","onlyCancer"),
c(t7,"onlyCancer","noCancer"),
c(t8,"onlyCancer","neither"),
c(t9,"Cancer","noCancer"),
c(t10,"Cancer","neither"))

colnames(CN_heat_cancerasd) = c("p value","phenotype1","phenotype2")
CN_heat_cancerasd$`p value` = as.numeric(CN_heat_cancerasd$`p value`)

```

# CN controlling age and sex
## onlyASD, onlyCancer, ASDCancer, neither
### colinearity - coorelation & vif
```{r}
for_CN$AGCNST = as.numeric(for_CN$AGCNST)

temp = for_CN
temp$pheno = as.numeric(as.factor(for_CN$phenotype))
temp$SEX = as.numeric(as.factor(for_CN$SEX))
GGally::ggpairs(dplyr::select(temp,SEX,pheno,AGCNST),
title = "Scatterplot Matrix for PHTS data")

car::vif(lm(CN ~ phenotype + SEX + AGCNST,data=for_CN))
car::vif(lm(burden ~ phenotype + haplogroup,data=for_CN))
```

###logistic regression - controlling age and sex
```{r}
temp = rbind.data.frame(onlyASD,onlyCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l1 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l2 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l3 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l4 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l5 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))

temp = rbind.data.frame(ASDCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l6 = summary(glm(phenotype ~ CN + SEX + AGCNST,data=temp, family = "binomial"))
```

```{r}
CN_heat = rbind.data.frame(
c(coef(l1)[2,4],"onlyASD","onlyCancer"),
#c(t2,"onlyASD","ASDCancer"),
c(coef(l2)[2,4],"neither","onlyCancer"),
#c(t4,"onlyASD","neither"),
#c(t5,"onlyCancer","neither"),
c(coef(l3)[2,4],"ASDCancer","onlyCancer"),
#c(t1,"onlyCancer","onlyASD"),
c(coef(l4)[2,4],"neither","onlyASD"),
#c(t3,"onlyCancer","ASDCancer"),
c(coef(l5)[2,4],"ASDCancer","onlyASD"),
c(coef(l6)[2,4],"ASDCancer","neither"),
#c(t6,"neither","ASDCancer"),
c(1,"ASDCancer","ASDCancer"),
c(1,"onlyASD","onlyASD"),
c(1,"onlyCancer","onlyCancer"),
c(1,"neither","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)

```

###logistic regression - controlling and sex
```{r}
temp = rbind.data.frame(onlyASD,onlyCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l1 = summary(glm(phenotype ~ CN + SEX,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l2 = summary(glm(phenotype ~ CN + SEX ,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l3 = summary(glm(phenotype ~ CN + SEX ,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l4 = summary(glm(phenotype ~ CN + SEX ,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l5 = summary(glm(phenotype ~ CN + SEX,data=temp, family = "binomial"))

temp = rbind.data.frame(ASDCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l6 = summary(glm(phenotype ~ CN + SEX, data=temp, family = "binomial"))
```

```{r}
CN_heat = rbind.data.frame(
c(coef(l1)[2,4],"onlyASD","onlyCancer"),
#c(t2,"onlyASD","ASDCancer"),
c(coef(l2)[2,4],"neither","onlyCancer"),
#c(t4,"onlyASD","neither"),
#c(t5,"onlyCancer","neither"),
c(coef(l3)[2,4],"ASDCancer","onlyCancer"),
#c(t1,"onlyCancer","onlyASD"),
c(coef(l4)[2,4],"neither","onlyASD"),
#c(t3,"onlyCancer","ASDCancer"),
c(coef(l5)[2,4],"ASDCancer","onlyASD"),
c(coef(l6)[2,4],"ASDCancer","neither"),
#c(t6,"neither","ASDCancer"),
c(1,"ASDCancer","ASDCancer"),
c(1,"onlyASD","onlyASD"),
c(1,"onlyCancer","onlyCancer"),
c(1,"neither","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)
```


###logistic regression - controlling age
```{r}
temp = rbind.data.frame(onlyASD,onlyCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l1 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l2 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyCancer,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l3 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l4 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))


temp = rbind.data.frame(onlyASD,ASDCancer)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l5 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))

temp = rbind.data.frame(ASDCancer,neither)
temp$phenotype = as.factor(temp$phenotype)
temp$SEX = as.factor(temp$SEX)
temp$CN = as.numeric(temp$CN)
temp$AGCNST = as.numeric(temp$AGCNST)
l6 = summary(glm(phenotype ~ CN + AGCNST,data=temp, family = "binomial"))
```

```{r}
CN_heat = rbind.data.frame(
c(coef(l1)[2,4],"onlyASD","onlyCancer"),
#c(t2,"onlyASD","ASDCancer"),
c(coef(l2)[2,4],"neither","onlyCancer"),
#c(t4,"onlyASD","neither"),
#c(t5,"onlyCancer","neither"),
c(coef(l3)[2,4],"ASDCancer","onlyCancer"),
#c(t1,"onlyCancer","onlyASD"),
c(coef(l4)[2,4],"neither","onlyASD"),
#c(t3,"onlyCancer","ASDCancer"),
c(coef(l5)[2,4],"ASDCancer","onlyASD"),
c(coef(l6)[2,4],"ASDCancer","neither"),
#c(t6,"neither","ASDCancer"),
c(1,"ASDCancer","ASDCancer"),
c(1,"onlyASD","onlyASD"),
c(1,"onlyCancer","onlyCancer"),
c(1,"neither","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)
```


## ASD vs. noASD 
### logistic regression
```{r}
for_CN_ASD$asd_phenotype = as.character(for_CN_ASD$asd_phenotype)
temp = for_CN_ASD
temp$pheno = as.numeric(as.factor(temp$asd_phenotype))
temp$SEX = as.numeric(as.factor(temp$SEX))
GGally::ggpairs(dplyr::select(temp,SEX,pheno,AGCNST),
title = "Scatterplot Matrix for PHTS data")

car::vif(lm(CN ~ phenotype + SEX + AGCNST,data=temp))
car::vif(lm(burden ~ phenotype + haplogroup,data=temp))

temp$SEX = as.factor(temp$SEX)
temp$asd_phenotype = as.factor(temp$asd_phenotype)
summary(glm(asd_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))


summary(glm(asd_phenotype ~ CN + SEX,data=temp,family = "binomial"))
summary(glm(asd_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))
```

## ASD vs. neither
### logistic
```{r}
temp1 = for_CN[which(for_CN$phenotype=="neither"),]
temp1$asd_phenotype = "neither"
temp = rbind.data.frame(ASD[,1:20],temp1)
temp$AGCNST = as.numeric(temp$AGCNST)
temp$asd_phenotype = as.factor(temp$asd_phenotype)

summary(glm(asd_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))
summary(glm(asd_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))
summary(glm(asd_phenotype ~ CN + SEX,data=temp,family = "binomial"))
```


## Cancer va. noCancer
### logistic
no sig
```{r}
for_CN_cancer$asd_phenotype = as.character(for_CN_cancer$asd_phenotype)

temp = for_CN_cancer
temp$pheno = as.numeric(as.factor(temp$asd_phenotype))
temp$SEX = as.numeric(as.factor(temp$SEX))
GGally::ggpairs(dplyr::select(temp,SEX,pheno,AGCNST),
title = "Scatterplot Matrix for PHTS data")

car::vif(lm(CN ~ phenotype + SEX + AGCNST,data=temp))
car::vif(lm(burden ~ phenotype + haplogroup,data=temp))
```
```{r}
temp$SEX = as.factor(temp$SEX)
temp$cancer_phenotype = as.factor(temp$cancer_phenotype)
summary(glm(cancer_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))


summary(glm(cancer_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))
summary(glm(cancer_phenotype ~ CN + SEX,data=temp,family = "binomial"))
```
## Cancer vs. neither
### logistic
```{r}
temp1 = for_CN[which(for_CN$phenotype=="neither"),]
temp1$cancer_phenotype = "neither"
Cancer = for_CN_cancer[which(for_CN_cancer$cancer_phenotype=="Cancer"),]
temp = rbind.data.frame(Cancer[c(1:19,21)],temp1)
temp$AGCNST = as.numeric(temp$AGCNST)


summary(lm(CN ~ cancer_phenotype + SEX + AGCNST,data=temp))
summary(lm(CN ~ cancer_phenotype + AGCNST,data=temp))
summary(lm(CN ~ cancer_phenotype + SEX,data=temp))
```
