---
title: "PHTS_mt_variant"
author: "R.W."
output: html_document
---

# knit setting
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library
```{r}
library(ggplot2)
library(ggsignif)
library(ggbeeswarm)
library(ggpubr)
library(beeswarm)
library(vipor)
library(ggfortify)
library(VennDiagram)
library(RColorBrewer)
library(corrplot)
library(factoextra)
library(epitools)
library(gridExtra)
```

# read in all polished data
```{r}
all = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_all.txt",sep="\t")
short_phenotype = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_short_phenotypes_wCN.csv",sep="\t")
short_all = read.csv("~/Documents/08PHTS_WGS_mt/variant/proband_remove/new_short_all.csv",sep="\t")

onlyASD = short_all[which(short_all$phenotype=="ASD"),]
onlyCancer = short_all[which(short_all$phenotype=="Cancer"),]
ASDCancer = short_all[which(short_all$phenotype=="ASDCancer"),]
neither = short_all[which(short_all$phenotype=="neither"),]

for_CN=rbind.data.frame(onlyASD,onlyCancer,ASDCancer,neither)
for_CN$AGCNST = as.numeric(for_CN$AGCNST)

for_CN_H = for_CN[which(for_CN$haplogroup=="H"),]
table(for_CN_H$pheno)
```

# wilcox test
## onlyASD, onlyCancer, neither, ASDCancer
```{r}
t1 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="ASD"),]$CN,for_CN_H[which(for_CN_H$phenotype=="Cancer"),]$CN)[3]
t2 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="ASD"),]$CN,for_CN_H[which(for_CN_H$phenotype=="ASDCancer"),]$CN)[3]
t3 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="ASDCancer"),]$CN,for_CN_H[which(for_CN_H$phenotype=="Cancer"),]$CN)[3]
t4 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="ASD"),]$CN,for_CN_H[which(for_CN_H$phenotype=="neither"),]$CN)[3]
t5 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="Cancer"),]$CN,for_CN_H[which(for_CN_H$phenotype=="neither"),]$CN)[3]
t6 = wilcox.test(for_CN_H[which(for_CN_H$phenotype=="ASDCancer"),]$CN,for_CN_H[which(for_CN_H$phenotype=="neither"),]$CN)[3]

CN_heat = rbind.data.frame(
c(t1,"onlyASD","onlyCancer"),
#c(t2,"onlyASD","ASDCancer"),
c(t3,"ASDCancer","onlyCancer"),
#c(t4,"onlyASD","neither"),
#c(t5,"onlyCancer","neither"),
c(t6,"ASDCancer","neither"),
#c(t1,"onlyCancer","onlyASD"),
c(t2,"ASDCancer","onlyASD"),
#c(t3,"onlyCancer","ASDCancer"),
c(t4,"neither","onlyASD"),
c(t5,"neither","onlyCancer"),
#c(t6,"neither","ASDCancer"),
c(1,"ASDCancer","ASDCancer"),
c(1,"onlyASD","onlyASD"),
c(1,"onlyCancer","onlyCancer"),
c(1,"neither","neither")
)

colnames(CN_heat) = c("p value","phenotype1","phenotype2")
CN_heat$`p value` = as.numeric(CN_heat$`p value`)
```

## wilcoxon compaerisions
```{r}
ASD = for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="ASD"),]
noASD = for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="noASD"),]
neither = for_CN_H[which(for_CN_H$phenotype=="neither"),]
Cancer = for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="Cancer"),]
noCancer = for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="noCancer"),]

wilcox.test(ASD$CN,noASD$CN)

wilcox.test(Cancer$CN,noCancer$CN)

wilcox.test(ASD$CN,neither$CN)

wilcox.test(Cancer$CN,neither$CN)

wilcox.test(Cancer$CN,ASD$CN)
```

## ASD, no ASD, neither
w/o covariants wilcox test
```{r}
t1 = wilcox.test(for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="ASD"),"CN"],for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="noASD"),"CN"])[3]
t2 = wilcox.test(for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="ASD"),"CN"],for_CN_ASD_H[which(for_CN_ASD_H$phenotype=="neither"),"CN"])[3]
t3 = wilcox.test(for_CN_ASD_H[which(for_CN_ASD_H$phenotype=="neither"),"CN"],for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="noASD"),"CN"])[3]


CN_heat_ASD_H = rbind.data.frame(
c(t1,"ASD","noASD"),
c(t2,"ASD","neither"),
c(t3,"neither","noASD"),
c(1,"ASD","ASD"),
c(1,"noASD","noASD"),
c(1,"neither","neither"))

colnames(CN_heat_ASD_H) = c("p value","phenotype1","phenotype2")
CN_heat_ASD_H$`p value` = as.numeric(CN_heat_ASD_H$`p value`)
```


## cancer, no cancer, neither
w/o covariants wilcox test
```{r}
t1 = wilcox.test(for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="Cancer"),"CN"],for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="noCancer"),"CN"])[3]
t1
t2 = wilcox.test(for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="Cancer"),"CN"],for_CN_cancer_H[which(for_CN_cancer_H$phenotype=="neither"),"CN"])[3]
t2
t3 = wilcox.test(for_CN_cancer_H[which(for_CN_cancer_H$phenotype=="neither"),"CN"],for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="noCancer"),"CN"])[3]
```

# co linearity 
```{r}
temp = for_CN_H
temp$pheno = as.numeric(as.factor(temp$phenotype))
temp$SEX = as.numeric(as.factor(temp$SEX))
GGally::ggpairs(dplyr::select(temp,SEX,pheno,AGCNST),
title = "Scatterplot Matrix for PHTS data")

car::vif(lm(CN ~ phenotype + SEX + AGCNST,data=temp))
car::vif(lm(burden ~ phenotype + haplogroup,data=for_CN_H))
```

# logitic regression
## ASD noASD 
```{r}
for_CN_ASD_H$asd_phenotype = as.character(for_CN_ASD_H$asd_phenotype)
temp = for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype!="neither"),]
temp$pheno = as.numeric(as.factor(temp$asd_phenotype))
temp$SEX = as.numeric(as.factor(temp$SEX))

temp$asd_phenotype = as.factor(temp$asd_phenotype)
temp$AGCNST = as.numeric(temp$AGCNST)

summary(glm(asd_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))

summary(glm(asd_phenotype ~ CN + SEX,data=temp,family = "binomial"))

summary(glm(asd_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))
```

## ASD vs. neither
```{r}
temp1 = for_CN_ASD_H[which(for_CN_ASD_H$phenotype=="neither"),]
temp1$asdneither_phenotype = "neither"
temp = for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="ASD"),]
temp$asdneither_phenotype = "ASD"
temp = rbind.data.frame(temp1,temp)

temp$SEX = as.factor(temp$SEX)
temp$asdneither_phenotype = as.factor(temp$asdneither_phenotype)
temp$AGCNST = as.numeric(temp$AGCNST)

summary(glm(asdneither_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))

summary(glm(asdneither_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))

summary(glm(asdneither_phenotype ~ CN + SEX,data=temp,family = "binomial"))
```

## Cancer noCancer - logistic regression
```{r}
for_CN_cancer_H$cancer_phenotype = as.character(for_CN_cancer_H$cancer_phenotype)
temp = for_CN_cancer
temp$cancer_phenotype = as.factor(temp$cancer_phenotype)
temp$SEX = as.factor(temp$SEX)
temp$AGCNST = as.numeric(temp$AGCNST)


summary(glm(cancer_phenotype ~ CN + SEX + AGCNST,data=temp,family = "binomial"))
summary(glm(cancer_phenotype ~ CN + SEX,data=temp,family = "binomial"))
summary(glm(cancer_phenotype ~ CN + AGCNST,data=temp,family = "binomial"))
```

## Cancer vs. neither
```{r}
temp1 = for_CN_H[which(for_CN_H$phenotype=="neither"),]
temp1$cancerneither_phenotype = "neither"
temp = for_CN_ASD_H[,c(1:19)]
temp$cancerneither_phenotype = "Cancer"
temp = rbind.data.frame(temp,temp1)
temp$AGCNST = as.numeric(temp$AGCNST)


summary(lm(CN ~ cancerneither_phenotype + SEX + AGCNST,data=temp))
summary(lm(CN ~ cancerneither_phenotype + AGCNST,data=temp))
summary(lm(CN ~ cancerneither_phenotype + SEX,data=temp))
```

# beeswarm plot for binary comparision results
```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="ASD" |for_CN_H$phenotype=="Cancer"),]
p_onlyasd_onlycancer = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("aquamarine","pink"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="ASD" |for_CN_H$phenotype=="ASDCancer"),]
p_onlyasd_asdcancer = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("pink","darkslateblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="ASD" |for_CN_H$phenotype=="neither"),]
temp$phenotype = factor(temp$phenotype,levels = c("ASD","neither"))
p_onlyasd_neither = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("pink","lightblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="Cancer" |for_CN_H$phenotype=="ASDCancer"),]
p_onlycancer_asdcancer = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("aquamarine","darkslateblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="Cancer" |for_CN_H$phenotype=="neither"),]
temp$phenotype = factor(temp$phenotype,levels = c("Cancer","neither"))
p_onlycancer_neitehr = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("aquamarine","lightblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="ASDCancer" |for_CN_H$phenotype=="neither"),]
temp$phenotype = factor(temp$phenotype,levels = c("ASDCancer","neither"))
p_neither_asdcancer = 
  ggplot(data=temp,aes(x=phenotype,y=CN)) +
  geom_boxplot(fill=c("darkslateblue","lightblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
p_asd = ggplot(data=for_CN_ASD_H,aes(x=asd_phenotype,y=CN)) +
  geom_boxplot(fill=c("lightcoral","gold"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  #scale_fill_brewer(palette="YlOrRd",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
p_cancer = ggplot(data=for_CN_cancer_H,aes(x=cancer_phenotype,y=CN,fill=cancer_phenotype)) +
  geom_boxplot(fill=c("darkgreen","greenyellow"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  scale_fill_brewer(palette="PuBu",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp = for_CN_H[which(for_CN_H$phenotype=="ASD"|for_CN_H$phenotype=="Cancer"),]
p_cancerasd = ggplot(data=temp,aes(x=phenotype,y=CN,fill=phenotype)) +
  geom_boxplot(fill=c("lightcoral","darkgreen"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  scale_fill_brewer(palette="PuBu",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp1 = for_CN_ASD_H[which(for_CN_ASD_H$asd_phenotype=="ASD"),]
temp1$temp = "ASD"
temp2 = for_CN_ASD_H[which(for_CN_ASD_H$phenotype=="neither"),]
temp2$temp = "neither"
temp = rbind.data.frame(temp1,temp2)
p_asdneither = ggplot(data=temp,aes(x=temp,y=CN,fill=temp)) +
  geom_boxplot(fill=c("lightcoral","lightblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  scale_fill_brewer(palette="PuBu",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
temp1= for_CN_cancer_H[which(for_CN_cancer_H$cancer_phenotype=="Cancer"),]
temp1$temp = "Cancer"
temp2= for_CN_cancer_H[which(for_CN_cancer_H$phenotype=="neither"),]
temp2$temp = "neither"
temp = rbind.data.frame(temp1,temp2)
p_cancerneither = ggplot(data=temp,aes(x=temp,y=CN,fill=temp)) +
  geom_boxplot(fill=c("darkgreen","lightblue"),alpha=0.8) +
  geom_quasirandom(dodge.width=0.9,alpha=.4) +
  theme_bw() +
  scale_fill_brewer(palette="PuBu",direction = 1) +
  ylab("Copy Number") +
  xlab("") +
  ylim(0,500) + 
  theme(axis.text.x=element_text(size=13,angle=0,hjust=0.5),
        axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15),
        #legend.text= element_text(size=13),
        #legend.title = element_text(""),
        legend.position = "none",
        title = element_text(size=20))
```

```{r}
p = arrangeGrob(p_onlyasd_onlycancer,p_onlyasd_asdcancer,p_onlyasd_neither,p_onlycancer_neitehr,p_onlycancer_asdcancer,p_neither_asdcancer,p_asd, p_cancer, p_asdneither,p_cancerneither,nrow=5)
```





